Free Theorems for Nested Types

### 2020-06-24 free-theorems-nesteds/paper3.tex
added explanation for coc of Nat types

\begin{dfn}\label{def:set-sem}
%Let $\rho$ be a set environment.
The {\em set interpretation} $\setsem{\cdot} : \F \to [\setenv, \set]$
is defined by
\begin{align*}
  \setsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \setsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
%  \setsem{\Gamma;\emptyset \vdash v} \rho &= \rho v \mbox{
%    if } v \in \tvars^0 \\ 
  \setsem{\Gamma; \ol\gamma
    %\emptyset
    \vdash \Nat^{\ol{\alpha}}
    \,F\,G}\rho &= \{\eta : \lambda \ol{A}. \,\setsem{\Gamma;
    \ol{\alpha} \vdash
    F}\rho[\ol{\alpha := A}] 
      \Rightarrow \lambda \ol{A}.\,\setsem{\Gamma; \ol\gamma,
        \ol{\alpha} \vdash G}\rho[\ol{\alpha := A}] \\ 
      &\hspace{0.3in}|~\forall \overline{A}, \overline{B} :
      \set. \forall \overline{R : \rel(A, B)}.\\ 
      &\hspace{0.4in}(\eta_{\overline{A}}, \eta_{\overline{B}})
      : \relsem{\Gamma; \ol{\alpha} \vdash F}\Eq_{\rho}[\ol{\alpha := R}]
      \rightarrow \relsem{\Gamma;\ol\gamma, \ol{\alpha} \vdash
        G}\Eq_{\rho}[\ol{\alpha := R}] \} \\
  \setsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\setsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \setsem{\Gamma;\Phi \vdash F+G}\rho &=
  \setsem{\Gamma;\Phi \vdash F}\rho +
  \setsem{\Gamma;\Phi \vdash G}\rho\\
  \setsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \setsem{\Gamma;\Phi \vdash F}\rho \times
  \setsem{\Gamma;\Phi \vdash G}\rho\\ 
  \setsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\set_{H,\rho})\ol{\setsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\set_{H,\rho}\,F & = \lambda
  \ol{A}. \setsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\set_{H,\rho}\,\eta &= \lambda
  \ol{A}. \setsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}]
\end{align*}
\end{dfn}
%The interpretations in Definition~\ref{def:set-sem} respect weakening,
%i.e., a type and its weakenings have the same set interpretations.
%
%= \setsem{\Gamma \vdash \sigma} \rho \to
%\setsem{\Gamma \vdash \tau} \rho$, as expected {\color{blue} by IEL}.
If $\rho \in \setenv$ and $\vdash F$ then we write $\setsem{\vdash F}$
instead of $\setsem{\vdash F}\rho$ since the environment is
immaterial. The third clause of Definition~\ref{def:set-sem} does
indeed define a set: local finite presentability of $\set$ and
$\omega$-cocontinuity of $\setsem{\Gamma;\ol{\alpha} \vdash F}\rho$
ensure that $\{\eta : \setsem{\Gamma;\ol{\alpha} \vdash F}\rho
\Rightarrow \setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho\}$
(which contains $\setsem{\Gamma;\ol\gamma \vdash
  \Nat^{\ol{\alpha}}\,F\,G}\rho$) is a subset of
$\big\{({\setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho[\ol{\alpha
      := S}]})^{(\setsem{\Gamma;\ol{\alpha} \vdash F}\rho[\ol{\alpha
      := S}])}~\big|~ \ol{S} = (S_1,...,S_{|\ol{\alpha}|}), \mbox{ and
}$ $S_i \mbox{ is a finite set for } i =
1,...,|\ol{\alpha}|\big\}$. There are countably many choices of tuples
$\ol{S}$, each of which gives rise to a morphism from
${\setsem{\Gamma;\ol{\alpha} \vdash F}\rho[ \ol{\alpha := S}]}$ to
${\setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho[\ol{\alpha :=
      S}]}$. But there are only $\set$-many choices of morphisms
between any two objects since $\set$ is locally small.  That
$\setsem{\Gamma; \ol\gamma \vdash \Nat^\alpha F\,G}$ is
$\omega$-cocontinuous follows from $\omega$-cocontinuity of
$\setsem{\Gamma;\ol\alpha \vdash F}$ and
$\setsem{\Gamma;\ol\gamma,\ol\alpha \vdash G}$. Interpretations of
$\Nat$-types ensure that $\setsem{\Gamma \vdash F \to G}$ and
$\setsem{\Gamma \vdash \forall \alpha. F}$ are as expected in any
parametric model.

>>>>

\begin{dfn}\label{def:set-sem}
%Let $\rho$ be a set environment.
The {\em set interpretation} $\setsem{\cdot} : \F \to [\setenv, \set]$
is defined by
\begin{align*}
  \setsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \setsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
%  \setsem{\Gamma;\emptyset \vdash v} \rho &= \rho v \mbox{
%    if } v \in \tvars^0 \\ 
  \setsem{\Gamma; \ol\gamma
    %\emptyset
    \vdash \Nat^{\ol{\alpha}}
    \,F\,G}\rho &= \{\eta : \lambda \ol{A}. \,\setsem{\Gamma;
    \ol{\alpha} \vdash
    F}\rho[\ol{\alpha := A}] 
      \Rightarrow \lambda \ol{A}.\,\setsem{\Gamma; \ol\gamma,
        \ol{\alpha} \vdash G}\rho[\ol{\alpha := A}] \\ 
      &\hspace{0.3in}|~\forall \overline{A}, \overline{B} :
      \set. \forall \overline{R : \rel(A, B)}.\\ 
      &\hspace{0.4in}(\eta_{\overline{A}}, \eta_{\overline{B}})
      : \relsem{\Gamma; \ol{\alpha} \vdash F}\Eq_{\rho}[\ol{\alpha := R}]
      \rightarrow \relsem{\Gamma;\ol\gamma, \ol{\alpha} \vdash
        G}\Eq_{\rho}[\ol{\alpha := R}] \} \\
  \setsem{\Gamma;\Phi \vdash \phi\ol{F}}\rho &=
  (\rho\phi)\,\ol{\setsem{\Gamma;\Phi \vdash
    F}\rho}\\
  \setsem{\Gamma;\Phi \vdash F+G}\rho &=
  \setsem{\Gamma;\Phi \vdash F}\rho +
  \setsem{\Gamma;\Phi \vdash G}\rho\\
  \setsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \setsem{\Gamma;\Phi \vdash F}\rho \times
  \setsem{\Gamma;\Phi \vdash G}\rho\\ 
  \setsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho &= (\mu
    T^\set_{H,\rho})\ol{\setsem{\Gamma;\Phi,\ol\gamma \vdash G}\rho}\\
    \text{where } T^\set_{H,\rho}\,F & = \lambda
  \ol{A}. \setsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\rho[\phi :=  F][\ol{\alpha := A}]\\
  \text{and } T^\set_{H,\rho}\,\eta &= \lambda
  \ol{A}. \setsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash
    H}\id_\rho[\phi := \eta][\ol{\alpha := \id_{A}}]
\end{align*}
\end{dfn}
%The interpretations in Definition~\ref{def:set-sem} respect weakening,
%i.e., a type and its weakenings have the same set interpretations.
%
%= \setsem{\Gamma \vdash \sigma} \rho \to
%\setsem{\Gamma \vdash \tau} \rho$, as expected {\color{blue} by IEL}.
If $\rho \in \setenv$ and $\vdash F$ then we write $\setsem{\vdash F}$
instead of $\setsem{\vdash F}\rho$ since the environment is
immaterial. The third clause of Definition~\ref{def:set-sem} does
indeed define a set: local finite presentability of $\set$ and
$\omega$-cocontinuity of $\setsem{\Gamma;\ol{\alpha} \vdash F}\rho$
ensure that $\{\eta : \setsem{\Gamma;\ol{\alpha} \vdash F}\rho
\Rightarrow \setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho\}$
(which contains $\setsem{\Gamma;\ol\gamma \vdash
  \Nat^{\ol{\alpha}}\,F\,G}\rho$) is a subset of
$\big\{({\setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho[\ol{\alpha
      := S}]})^{(\setsem{\Gamma;\ol{\alpha} \vdash F}\rho[\ol{\alpha
      := S}])}~\big|~ \ol{S} = (S_1,...,S_{|\ol{\alpha}|}), \mbox{ and
}$ $S_i \mbox{ is a finite set for } i =
1,...,|\ol{\alpha}|\big\}$. There are countably many choices of tuples
$\ol{S}$, each of which gives rise to a morphism from
${\setsem{\Gamma;\ol{\alpha} \vdash F}\rho[ \ol{\alpha := S}]}$ to
${\setsem{\Gamma;\ol\gamma,\ol{\alpha} \vdash G}\rho[\ol{\alpha :=
      S}]}$. But there are only $\set$-many choices of morphisms
between any two objects since $\set$ is locally small.  That
$\setsem{\Gamma; \ol\gamma \vdash \Nat^\alpha F\,G}$ is
$\omega$-cocontinuous follows from $\omega$-cocontinuity of
$\setsem{\Gamma;\ol\alpha \vdash F}$,
$\setsem{\Gamma;\ol\gamma,\ol\alpha \vdash G}$, and the equality
relation environment $\Eq$. Interpretations of $\Nat$-types ensure
that $\setsem{\Gamma \vdash F \to G}$ and $\setsem{\Gamma \vdash
  \forall \alpha. F}$ are as expected in any parametric model.
